\documentclass[../main.tex]{subfiles}
\begin{document}
\lecture{22}{Fri 13 Jan}{Differentials}
\begin{lemma}
Let $f:X\to Y$ be a surjective morphism  over a scheme $S$.\\
Assume $\pi_X: X\to S$ is proper and $\pi_Y:Y \to S$ is separated, then $\pi_Y$ is proper.
\end{lemma}
\begin{proof}
Need to show $\pi_Y$ is universally closed.\\
Let $T\to S$ be any scheme and $Z \subset Y_T$ closed.\\
By some exercise, $f_T: X_T\to Y_T$ is surjective.\\
So $\pi_{Y,T} ( Z) = \pi_{X,T} ( f_T^{-1}( Z) ) $ and the lhs is closed.
\end{proof}
\begin{propo}
Let $X$ be a proper scheme over a field $k$.\\
Assume $X$ is geometrically raduced and geometrically connected.\\
Then $\Gamma( X,\O_X) = k$ and $\Gamma( X, \O_X^{\times}) = k^{\times}$ 
\end{propo}
\begin{proof}
By the above, $\Gamma( X_{\overline{K}} , \O_{X_{\overline{K}} } ) = \Gamma( X,\O_X) \otimes_K \overline{K}$.\\
So wlog, $K$ algebraic closed.\\
Consider any morphism $f:X\to \mathbb{A}^{1}_K$.\\
Since $X\to \spec K$ is proper and $\mathbb{A}^{1}_K \to \spec k$ is separated so $f( X) $ is closed.\\
We know $f( X) $ is not $\mathbb{A}^{1}_K$ since $ \mathbb{A}^{1}_K \to \spec K$ is not proper.\\
So $f( X) $ is closed, connected, reduced ie. a closed point.\\
In particular, $f( X) $ is a rational point.\\
This means that $\Gamma( X,\O_X) \simeq \hom_{ ( Sch /k) } ( X, \mathbb{A}^{1}_K) = \hom( \spec k , \mathbb{A}^{1}_K) = k$ 
\end{proof}
\begin{crly}
Let $X$ be a geometrically integral scheme, proper over a field $k$.\\
Then, for any effective cartier divisor  $D$ on $X$, there is a bijection
\[ 
|D| \leftrightarrow \frac{\Gamma( X,\O_X( D) ) \setminus \left\{ 0 \right\} }{k^{\times}}
\]

\end{crly}
\section{Differentials}
Recall that a local ring $A$ is regular if $\dim_{ A /m}  \frac{m}{m^{2}}= \dim A$.\\
We want to introduce a sheaf $\Omega$ such that $\Omega_X \otimes_{\O_{Y,x} } k( x) \simeq m /m^{2}$ 
\begin{defn}[Derivations]
	Let $A$ be a ring, $B$ an algebra over $A$ and $M$  a $B$-module.\\
	An $A$-linear derivation from $B$ to $M$ is a map $d:B \to M$ such that 
	\begin{itemize}
	\item $d( b+b') = db + db' \forall b,b' \in B$ 
	\item $d( bb') = db \cdot b' + b' \cdot db$ 
	\item $da = 0 \forall a \in A$
	\end{itemize}
\end{defn}
\begin{defn}
Let $A$ be a ring and $B$ an $A$-algebra, then the module of relative differentials forms (or Kaehler differentials) of $B$ over $A$ is the universal $B$-module $\Omega_{B / A} $ together with an $A$-linear derivation $d: B \to \Omega_{B / A} $ through which all derivations $d: B \to M$ factor.
\end{defn}
\begin{exemple}
Let $A$ be a ring and $B = A [ x_1,\ldots,x_n] $ then $\Omega_{B / A} $ is the free $B$-module generated by the $dx_i$ 
\end{exemple}
\begin{propo}
Let $A$  be a ring, $B,C$ $A$-algebras, $S \subset B$ mult. closed, $S' \subset A$ mult. closed mapping to units in $B$, then there are natural isoms.
\begin{align*}
\Omega_{B \otimes_A C / A } \simeq ( \Omega_{B / A} \otimes_A C) \oplus ( \Omega_{C / A} \otimes_A B) \\
\Omega_{S^{-1}B / A} \simeq S^{-1}\Omega_{B / A} \\
\Omega_{B / A} \simeq \Omega_{B / S'^{-1}A} 
\end{align*}
\end{propo}
Now, if $f:X\to Y$ is a morphism of schemes, $\spec A = V \subset Y$ open affine and $U = f^{-1}( V) = \bigcup \spec B_i$, then the $\widetilde{ \Omega_{B_i / A} }$ glue to a sheaf.
\begin{defn}
	Let $f:X\to Y$ be a morphism of schemes, the qcoh. sheaf $\Omega_{X /Y} $ is called the sheaf of relative differentials or cotangent sheaf.
\end{defn}	
\begin{propo}
Let $f:X\to Z$ be a morphism of schemes
\begin{enumerate}
\item Let $g:Y \to Z$ be another morphism of schemes, then there is a natural isom
	\[ 
	\Omega_{X \times_Z Y/ Z	} \simeq pr_1^{\ast}\Omega_{X /Z} \oplus pr_2^{\ast}\Omega_{Y /Z} 
	\]
\item Let $U \subset X$ open, then there is a natural isom
	$\Omega_{X / Y /}|_U \simeq \Omega_{ U /Y} $ 
\item Let $V \subset Y$ be an open subscheme, such that $f$ factors through $V$ then there is a natural isom $\Omega_{X /V} \simeq \Omega_{X/Y} $ 
\item Let $g:Y \to Z$ be a morphism of schemes, then there is a natural exact sequence of $\O_X$-modules
\[ 
f^{\ast}\Omega_{Y /Z} \to \Omega_{X /Z}  \to \Omega_{X /Y} \to 0
\]
\item Let $Z \subset X$ be a closed subscheme with ideal sheaf $ \mathcal{I}$, then there is a natural right exact sequence
	\[ 
	\mathcal{I} / \mathcal{I}^{2} \to \Omega_{X / Y} \otimes_{\O_X} \O_Z \to \Omega_{Z /Y} \to 0
	\]
	
\item Let $g:Y' \to Y$ be a morphism of schemes, then there is a natural isom.
	\[ 
	\Omega_{X_{Y'} / Y'} \simeq g'^{\ast}\Omega_{X /Y} 
	\]
	where $g': X_{Y'} \to X$ is the natural projection map.

\end{enumerate}
\end{propo}
\begin{crly}
Let $B$ be a localization of a f.g. $A$-algebra, then $\Omega_{B / A} $ is a f.g. $B$-module.\\
In particular, if $f:X\to Y$ is of finite type and $X$ is Noetherian, then $\Omega_{X /Y} $ is coherent.
\end{crly}
When is the conormal sequence left-exact?\\
\begin{propo}
Let $f:X\to Y$ be a morphism of schemes and $\iota:Z \to X$ be a closed subscheme with ideal sheaf $\mathcal{I}$.\\
Assume there is a left-inverse $X\to Z$, then the conormal sequence is exact.
\end{propo}




\end{document}	
