\documentclass[../main.tex]{subfiles}
\begin{document}
\lecture{20}{Mon 19 Dec}{principal divisors, class group}
Throughout, $X$ is integral, Noetherian and $R1$.\\
\begin{defn}[Principal divisors]
\begin{enumerate}
\item A weil divisor $D$ on $X$ is called principal if $D= div ( f) $ for some $f\in K( X)^{\times}$.
\item Two Weil divisors $D,D'$ on $X$ are called linearly equivalent if $D- D'$ is principal.
\item The ( Weil divisor) class group $Cl( X)  $ of $X$ is the quotient of $Div( X) $ by the subgroup of principal divisors.
\end{enumerate}
\end{defn}
\begin{lemma}
Assume $X$ is noetherian, integral, normal and let $s$ be a rational section of an invertible sheaf.\\
If $v_Z( s) \geq 0\forall Z \subset X$ prime, then $s$ extends uniquely to a global section of $ \mathcal{L}$.
\end{lemma}
\begin{propo}
Let $X$ be an integral, normal, Noetherian scheme, then $div$ is injective.
\end{propo}
\begin{proof}
Assume $div(  \mathcal{L},s) =0$. In particular $v_Z( s) =0\forall Z \subset X$ prime.\\
By the above, wlog, $s$ is a global section.\\
Consider $s\times - : \O_X \to \mathcal{L}$.\\
If this is an iso, then $( \O_X,1) \simeq ( \mathcal{L},s) $.\\
To check that $s\cdot$ is an iso, we can work locally, ie., we may assume $X= \spec A$ and $ \mathcal{L}= \O_X$ is trivial, $s\in A$.\\
Since $v_Z( s) =0 \forall Z \subset X$ prime, we have $v_Z( \frac{1}{s}) =0$ 
\end{proof}
\begin{defn}
	Let $D$ be a Weil divisor on $X$, the sheaf asscoiated to $D$ is 
	\[ 
	\O_X( D): U \to \left\{ f\in K( X)^{\times}|div|_U f + D|_U \geq 0 \right\} \cup \left\{ 0 \right\} 
	\]
	where $div|_U f$ is the principal divisor asscoiated to $f$ considered as a rational function on $U$.
Note that $1\in K( X) $ is a rational section of $\O_X( D) $.	
\end{defn}
This is a sheaf.
\begin{lemma}
$\O_X( D) $ is quasi-coherent.
\end{lemma}
\begin{proof}
By a theorem, it suffices to show that for all open affines $\spec A = U \subset X$ open and $a \in A \setminus \left\{ 0 \right\} $, the map
\[ 
\rho: \O_X( D) ( U)_a \to \O_X( D) ( D( a) ) 
\]
is an isomorphism.
\end{proof}
\begin{propo}
Let $X$ be a normal integral and Noetherian scheme.\\
Let $ ( \mathcal{L},s) $ be a pair of an invertible sheaf and a rational section $s$.\\
Then, there exists an isomorphism $\O_X( div( \mathcal{L},s) ) \to \mathcal{L}$ that sends $1$ to $s$.
\end{propo}
\begin{proof}
Let $\emptyset \neq U \subset X$ such that $ s\in \mathcal{L}( U) $.\\
Let $V \subset X$ be any open, let $t \in\O_X( div(  \mathcal{L},s) ) ( V) \subset K( X) $.\\
Then $t|_{V\cap U} \in \O_X( div(  \mathcal{L},s) ) ( V) \subset K( X) $ and
\[ 
div t + div( s) |_{U\cap V}  \geq 0
\]
Then $t|_{U\cap V} $ has no poles, so $t\in \O_X( V\cap U) $.\\
So $st \in \mathcal{L}( V\cap U) $ which we can extend to $ \mathcal{L}( V) $.\\
So we get a morphism $\O_X( div( s) ) \to \mathcal{L}$.
\end{proof}
\begin{defn}
	A weil divisord $D$ on $X$ is called locally principal if $\forall x \in X\exists x\in U \subset X$ such that $D|_U$ is principal.
\end{defn}
\begin{crly}
Let $D$ be a locally principal Weil divisor on a normal, integral, Noetherian scheme.\\
Then $\O_X( D) $ is invertible.
\end{crly}
\begin{proof}
	We have $\O_U( D|_U) = \O_X( D) |_U$.\\
	If $D|_U$ is principal, then $\O_U( D|_U) \simeq \O_U$ 
\end{proof}
\begin{propo}
Let $X$ be a normal, integral, Noetherian.\\
Let $D$ be a Weil divisor on $X$ such that $\O_X( D)$ is invertible, then $D= div( \O_X( D),1 ) $, in particular $D$ is locally principal.
\end{propo}
\begin{propo}
If $A$ is a Noetherian integral domain, tfae\\
$A$ is factorial $\iff$ every height 1 prime in $A$ is principal.
\end{propo}
\begin{propo}
Factorial domains are normal.
\end{propo}
\begin{thm}
	All Noetherian regular local rings are factorial.
\end{thm}

\begin{defn}[Factorial scheme]
	A scheme $X$ is called factorial if $\O_{X,x} $ is factorial $\forall x \in X$.
\end{defn}
\begin{propo}
Let $X$ be a factorial, integral, Noetherian scheme, then, every Weil divisor on $X$ is locally principal.\\
In particular $div$ is an isomorophism and $Pic( X) \sieq Cl( X) $
\end{propo}
\begin{proof}
	Wlog, $X= \spec A, D \subset X$ prime with ideal $p \subset A$.\\
	Note that $D|_{X\setminus D} = 0$.\\
	Let $x\in D \subset X$, then $ht( p A_x) =1$ and $A_x$ is factorial and $p A_x= ( f_x) $ for some $f_x \in A_x$.\\
	In particular, $v_D( f_x) = 1$, so $ div( \O_X, f_x) |_U = D|_U$ for some open neighborhood $x\in U \subset X$ 
\end{proof}
\begin{crly}
If $A$ is factorial, then $Pic( \spec A) = Cl ( \spec A) =0$ 
\end{crly}
\begin{exemple}
\begin{enumerate}
\item If $A$ is a Dedekind domain, then $Cl( A) $ are fractional ideals mod principal fractional ideals and this is isomorphic to $Pic( \spec A) $ 
\item If $k$ is a field, then
	\[ 
	Pic( \mathbb{A}^{n}_k ) = Cl( \mathbb{A}^{n}_K) =0
	\]
\end{enumerate}
\end{exemple}
\begin{propo}
Let $X$ be a normal, integral, Noetherian scheme and $Z \subset X$ is integral and closed, let $U = X \setminus Z$, then
\begin{enumerate}
\item The restriction map $Cl( X) \to Cl( U) $ is surjective
\item If $codim( Z,X) \geq 2$, then $Cl( X) \simeq Cl( U) $ 
\item If $codim( Z,X) =1$, then there is an exact sequence
	\[ 
	\mathbb{Z}\to Cl( X) \to Cl( U) \to 0
	\]
	
\end{enumerate}

\end{propo}




	



\end{document}	
