\documentclass[../main.tex]{subfiles}
\begin{document}
\lecture{2}{Tue 01 Mar}{Stuff}
There are 3 different possibilities for optimisation problems
\begin{enumerate}
\item The finite max exists
\item The problem is unbounded
\item The problem is infeasible, ie. $S = \emptyset$ 
\end{enumerate}
There is a special class of functions $f$ which allow us to solve problems rather easily
\begin{itemize}
\item $f$ is linear
\item $S$ is made up of linear inequality constraints.
\end{itemize}
Such an optimisation problem is called a linear program.
\subsection{Geometry of linear equalities}
What can the feasible region of a LP look like?\\
It will look like an intersection of half spaces, we call it a polyhedron.

\begin{propo}
Polyhedra are convex
\end{propo}
Indeed they are finite intersections of half-spaces, which are all convex.
Given vectors $v_1,\ldots, v_n$ we can look at
\begin{enumerate}
\item linear combinations ( span) 
\item positive linear combinations $\sum a_i v_i, a_i \geq 0$ ( cone) 
\item convex combinations $\sum a_i v_i, 0 \leq a_i \leq 1, \sum_i a_i =1$ ( convex hull) 
\end{enumerate}
If we have a linear program, only active constraints can be involved in optimal solutions.\\
Given a linear program, we want to turn it into a standard form.
\begin{defn}[Equivalent linear programs]
	Two linear programs $P $ and $P'$  are equivalent if for every feasible $x$ in $P$, there exists a feasible $x' \in P'$ with $value_P( x) = value_{P'} ( x') $ and conversely.\\
	Furthermore, finding $x$ from $x'$ and $x'$ from $x$ is easy.
\end{defn}
What does a general linear program look like?
\[ 
\max c \cdot x
\]
s.t. $Ax \geq b$.
\begin{propo}
every linear program can be transformed into a linear program in inequality standard form.
\end{propo}
Given an equality constraint $a\cdot x \geq b$, we can turn it into $a \cdot x =b+w$ with $w \geq 0$.\\
Such a variable $w$ is called a slack variable.\\
Hence, given the constraint $Ax \geq b$, we can turn it into an equality constraint
\[ 
	Ax = b + \left[ w_1, \ldots, w_n \right]^{T}
\]
and ask $w_i \geq 0\forall i$ 



\end{document}	
